Application Specific Arabic Morphological Analyzer for Arabic Text Mining

### 2011-02-21 trunk/tex/sarf.aclhlt.tex
made some structural changes, and added some needed sentences, although this did not reduce size, but helped creating a better structure of the paper, reduction should come next after we agree on what to keep and what to remove

%\title{Instructions for NAACL HLT 2010 Proceedings\Thanks{This...}}
\title{Case-based Arabic Morphological Analyzer for Arabic Text Mining}

@@

\begin{document}
\maketitle
\begin{abstract}
Natural language processing uses morphological 
analysis to abstract and annotate text.
Often times, 
%      and in particular in the context of the Arabic language, 
annotations resulting from 
a branch of morphological analysis may not be
appropriate for the case study at hand. 
In this paper we present Sarf, a case based morphological 
analyzer for Arabic. 
Sarf allows a case-based controller to control and refine
the state of the analysis on the fly. 
It uses non-deterministic finite state Automata
to analyze a stream of text. 
Each alive machine corresponds to a valid analysis. 
Sarf uses concatenative analysis based on recursive affixes 
to better preserve part of speech information.
We automated the analysis of three books of Islamic
narrations (hadith) using Sarf where we
abstracted each book into a vector of complex structures.
Our results show higher accuracy and better efficiency
compared to current analyzers.
\end{quote}
\end{abstract}

@@

%Automated analysis of Arabic data sets, including texts, 
%publications, records and digital media is essential
%with the huge digital Arabic content available nowadays. 
Morphological analysis is key in current automated 
analysis techniques for Arabic text. 
Current morphological analyzers~\cite{Sughaiyer:04}
use concatenative analysis
consider the internal structure 
of an Arabic word and
compose it into several {\em morphemes}. 
A morpheme can be a {\em stem}, or an {\em affix}.
An affix can be a {\em prefix, suffix, } or an {\em infix}.
The analysis of one word may lead to several possible
morphological solutions.
\vocalize
For instance, the word \RL{'a.hmadH}~\footnote{In this document, we use the default ArabTeX transliteration style ZDMG.}
may have two valid morphological analyses. 
The letter \RL{'a} may be a prefix and the word means 
``I praise him'', or 
%The letter \RL{'a} may also be 
part of the stem \RL{'a.hmad} (a proper noun)
and the word means ``his Ahmad''.

\novocalize
The accuracy of the solutions suffer due to inherent difficulties
of morphological analysis of the Arabic language. 
For example, it is common practice to write Arabic text
without short vowels. 
This greatly increases the ambiguity of Arabic text. 
Arabic letters can have up to 
four different forms
corresponding to their position in a word, i.e, beginning,
middle, end of word and separate forms. 
This allows the phrase \transfalse
\RL{il_A\nospace almdrsT} \transtrue
to be visually recognizable
as two separate words \RL{il_A} (to) and \RL{almdrsT} (the school) 
without the need of a space in between. 
The reason is the first word \RL{il_A} ends with
\RL{_A} a non-connecting letter. 
These words,
referred to as ``run-on'' words~\cite{Buckwalter:04},
occur often, and greatly increase the
difficulty of tokenization.

Current morphological analyzers such as 
Buckwalter~\shortcite{Buckwalter:02},
Beesley~\shortcite{Beesley:01},
SAMA~\cite{Kulick:10},
and ElixirFM~\cite{Otakar:07} 
take as input white space delimited tokens~\cite{Kulick:10},
consider them as words,
and enumerate all possible solutions. 
This approach has several problems. 
First, a white space delimited token may have 
more than one word.
Second, the exhaustive enumeration may hurt performance and may
not be necessary or appropriate
in some case studies as noted in~\cite{Maamouri:10}. 
Other morphological analyzers such as 
Amira~\cite{Diab:07,Benajiba:07},
MAGEAD~\cite{Habash:05}, and MADA+TOKAN~\cite{Habash:09} 
use machine learning and support vector machines (SVM) 
to enhance the accuracy of the morphological analysis at the expense 
of performance.
Xerox rules can be compiled into specialized finite state
machine (FSM) based analyzers as described in~\cite{Beesley:03}.
The efficiency of the resulting analyzer depends on the
way the Xerox rules are written. 
This requires deep knowledge and insight from the user
in compilation techniques, context free grammars, 
and morphological analysis.
%context free grammar rules that can be automatically compiled
%into efficient non-deterministic finite state Automata (NDFSA).

We hypothesize that many NLP case studies need the 
morphological analyzer to answer queries that do not need 
high accuracy at the low morphological analysis level.
The analysis at the higher levels can often compensate for 
tolerable inaccuracy at low levels. 
For example, if a query concerns proper names and the 
analyzer is considering a 
prefix that connects only to a verb,
the analyzer can answer without considering the 
rest of the word.

We find evidence to our hypothesis in~\cite{Maamouri:10}.
The addition of a new corpus to the Arabic Tree Bank 
(ATB)~\cite{Maamouri:04}
required compliance with the DARPA GALE program 
including several NLP tasks such as English translation 
annotations and word-level alignment of Arabic and English. 
The new data challenged the existing processing and 
annotation guidelines. 
This lef to refining 
SAMA~\cite{Kulick:10} with 
alternative guidelines where the NLP 
task interacted with the morphological analyzer.

The work of Habash~\shortcite{Habash:06} reports that different 
types of morphological analyses behave differently on the same case 
study. 
The analyzers may find different solutions or 
may present the solutions in different orders. 
The presented solutions and the order of presentation may not be 
appropriate for the case study. 
We find strong evidence in our results 
that a case-based morphological analyzer that allows a case-based 
controller to intervene at every decision to guide, use, order, and 
refine the morphological analysis is of high utility.

%\subsection{Sarf}
%\label{sec:intro:sarf}

{\bf Sarf.~~~}
In this paper, we present Sarf~\footnote{Sarf (name modified for blind review) 
will be available online as an open source tool.},
 a {\em novel case-based efficient
morphological analyzer} that uses 
non-deterministic finite state Automata 
driven by a case-based controller.
Each alive machine in Sarf represents a valid morphological analysis
and takes as input one letter at a time.
The case-based controller prunes false positives
early in the run by making a decision at every input letter
to control the alive machines.
%Each alive machine in Sarf takes as input one letter at a time 
%from a text stream. 
Sarf does not assume that the word at hand is a token and
performs tokenization on the fly based on morphological correctness.
Up to our knowledge, this is the first morphological analyzer that 
handles the ``run-on'' words problem.

Sarf uses concatenative analysis with a novel 
{\em recursive affixes} refinement. 
Sarf introduces recursive affixes in order to
better retain part of speech information and enhance the 
efficiency of affix matching.

\transfalse
\begin{figure}[tb]
\center{
\resizebox{.9\columnwidth}{!}
{ \input{figs/exhadith.pdftex_t}}
\caption{Hadith abstraction example.}
\label{f:exhadith}
}
\end{figure}
% the FSMs for the three words
\transtrue

We validate our hypothesis using a case study from the Islamic 
literature. 
A \RL{.hady_t} is a narration related to the prophet Mohammad
through a \RL{sanad} or a sequence of narrators. 
Figure~\ref{f:exhadith} shows an example \noArRL{.hady_t} in 
Arabic with its transliteration and translation. 
The authenticity of a \noArRL{.hady_t} depends on 
the credibility of the narrators as reported in 
separate biography books. 
In this paper, we consider the problem of automatically segmenting
a \noArRL{.hady_t} book into narrations, each 
narration into its \RL{matn} (content) and its \noArRL{sanad},
and each sanad into the 
separate narrators. Now, we can look each narrator 
up in the biography books. 
For example, 
Figure~\ref{f:exhadith} shows proper names in boxes connected
to form complex names of narrators. 
For example, 
\noTrRL{qtybT} is the first name 
of narrator $n_1$, and 
\noTrRL{s`yd} is the name of his father as 
the word \noTrRL{bn} (son of) indicates.

The hadith case study is interesting to Sarf since the controller
considers %most of the analyses where we have a concentration
proper names and ignores text where
proper names occur rarely.% as long %as the text is valid.
We are also interested in discovering a words 
that relate names to each other such as \RL{bin} or ``the son of''
and words that mean narrate such as \RL{qaal} or ``said''.

With Sarf, we successfully automated the analysis of 
three books of Islamic narrations~\cite{IbnHanbal,AlTousi,AlKulayni}
and extracted from each one of them a vector of complex structures 
that are three level deep in hierarchy.
%We also showcase that the results of the accurate high level 
%abstraction
%results of the case-based analysis was not affected by a major
%refinement to the analyzer that increased low level analysis accuracy. 
We make the following key contributions. 
\begin{enumerate}
\item {\bf Case-based analysis:}  We designed and implemented
a case-based morphological analyzer where a controller
guides the analysis. 
%\item {\bf Exhaustive analysis:} Since we have a case-based 
%controller that can decide on the fly whether an analysis is 
%accepted or not, we can afford to keep all valid analyses modulo
%those pruned by the case controller. 
%We do that using non-deterministic FSAs. 
\item {\bf Recursive affixes:}
We refine the concatenative analysis of 
Buckwalter~\shortcite{Buckwalter:02} and define affixes
recursively. 
This allows Sarf to perform better and 
maintain compositional part of speech (POS) tags.
\item {\bf Hadith case study:}
%We evaluated Sarf using the hadith case study. 
We were able to efficiently extract a complex three 
level hierarchical 
structure containing the chains of narrators 
with high accuracy. 
\end{enumerate}

@@

The rest of this paper is structured as follows. In Section~\ref{sec:related}
we compare Sarf to related work. In Section~\ref{sec:sarf}
we discuss Sarf and illustrate it with a running example.
%We also explain how non-deterministic finite state Automata
%work. 
In Section~\ref{sec:islamic} we present the hadith case
study. We present our results in Section~\ref{sec:results}
and conclude with future work in Section~\ref{sec:future}.

@@

The survey in~\cite{Sughaiyer:04} compares
several morphological analyzers. 
Analyzers such as~\cite{Khoja:01,Darwish:02} 
target specific applications in the 
analyzer itself or use a specific set of POS tags
as their reference.
Sarf differs in that it is a general morphological 
analyzer that reports all possible solutions. 
It is case-based in the sense that a controller 
prunes these solutions.

Sarf builds upon the lexicon of Buckwalter\shortcite{Buckwalter:02}.
It differs from Buckwalter in that it defines a shorter list of affixes
and a longer list of affix compatibility rules to allow recursive 
affixes so that we can have prefix-prefix an suffix-suffix 
concatenations.
This allows Sarf to better maintain and propagate 
the POS tags associated with affixes. 
%This information is highly 
%redundant with Buckwalter.% and that may lead to consistency issues. 
Sarf keeps all the possible analyses alive where each analysis
corresponds to a set of decisions and positions in the text. 
Buckwalter produces a set of solutions for each token 
and the several analysis threads of the text need to be 
generated as a product of the token solutions.

Similar to Sarf, Beesley~\shortcite{Beesley:01} uses
finite state Automata (FSA). 
It reports that the number of Automata generated by a compiler
for Xerox rules can not be controlled and also reports a difficulty 
to compose the FSAs into a single framework. 
Sarf constructs an NDFSA from
the controller and the prefix, stem, and suffix deterministic 
transducers to solve the problem.
Sarf differs in that it uses recursive affixes in the 
affix transducers, and in that the transducers consider
input from the controller to modify their behavior. 
%With Sarf it is easy to control the FSA with a case based
%controller. 
Beesley requires recompilation.% of the Xerox rules.

@@

Our approach is close to the refinement of SAMA~\cite{Maamouri:10}.
SAMA was refined to interact with
the ATB~\cite{Maamouri:04} project after the addition of a large 
new corpus. 
%We find supporting evidence in this decision to our hypothesis
%about the utility of a case-based analyzer. 
The algorithmic changes in SAMA were
done manually and worked in integration with the ATB format. 
Sarf is a case-based controller approach that can modify 
the behavior of the analyzer on the fly and can interact
with any case study. 
%In Sarf, the analyzer remains general, and the 
%case-based controller specializes the analysis.

@@

Like ElixirFM~\cite{Otakar:07}, Sarf builds on the lexicon
of the Buckwalter analyzer. 
Sarf also uses deterministic parsing with tries 
to implement the affix and stem transducers. 
We think that the inferential-realizational approach 
of ElixirFM
that is highly compatible with the Arabic linguistic 
description~\cite{Badawi:04}
can benefit from many features unique to the Arabic language.
Sarf leaves implementing that to the case-based controller
since in many cases the abstraction needs only a partial 
linguistic model of the Arabic language. 
%We can compare the case-based controller in Sarf to a high level
%component in the ElixirFM domain specific language.

@@

Before moving with the letter \RL{b} to the accept state,
$\Phi$ needs to make sure that the stem \RL{l`b} is compatible
with the prefix \RL{wsy}. 
Sarf keeps a compatibility category values as part
of the accept states. 
%Sarf considers the value of the category as part of the state.
Thus each accecpt state in Figure~\ref{f:example} represents
more than one concrete state. 
If the category of \RL{l`b} is compatible with the category of
\RL{wsy} then \RL{b} will move $\Phi$ to an accept state. 
Otherwise, it will move it to a regular or a reject state.

Since $\Phi$ is now in an accept state in ${\cal S}$, it will
spawn $\Xi$. $\Xi$ transitions to the root of ${\cal X}$. 
$\Phi$ die when it consumes \RL{h} since
no edge from the current state is labeled with \RL{h}. 
$\Xi$ represents a valid analysis and can proceed.

@@

A hadith in Islamic literature is a narration from the prophet Mohammad
related by multiple narrators.
Establishing the authenticity of a hadith is an important task
in Islamic studies. 
Based on that Islamic scholars issue rules that affect the life
of people around the world. 
The literature lacks an exhaustive mechanism of authentication
checks. This is currently manual work prone to error and almost
impossible to complete in the life time of a scholar due to the
large literature available. 
%Several researchers~\cite{Hadithopaedia:08} attempted to automate 
%the analysis of the hadith literature. 
In this paper, we use Sarf to successfully automate the
analysis of three books of hadith selected 
arbitrarily~\cite{IbnHanbal,AlKulayni,AlTousi}\footnote{We obtained
  the digitized books from online sources such as 
  \href{http://www.yasoob.com/}{http://www.yasoob.com/} and 
  \href{http://www.al-eman.com/}{http://www.al-eman.com/}. }
%  Our understanding is that those books are historical documents with 
 % public IP.}.

@@

We investigated the case of Al-Kafi where the accuracy of 
the simple analysis was significantly lower than that of the 
refined analysis. 
We discovered that the chains detected by the refined analysis 
were a subset of those detected by the simple
analysis. 
The reason for the lower accuracy is false positive chains, 
implying that the simple analysis did not miss important 
information.

We hypothesize that a higher level of abstraction such as 
intersecting the narrators from the hadith books with the narrators 
in the biographies can reduce the false positives in the chains
in this case.

\subsection{Buckwalter and ElixirFM}

We implemented our case study with Buckwalter in two phases. 
In phase one, we processed the entire text with Buckwalter and
obtained annotations to text. 
In the second phase, we fed the annotations to 
the hadith controller as if they were coming from Sarf. 
We measured the time it took to accomplish both tasks
less the time spent streaming the annotations.
We did the same for ElixirFM.

Sarf outperforms both Buckwalter and ElixirFM in terms of 
running time. 
It also produces higher accuracy measures 
mainly due to the extended
lexicon as well as to the fact that Sarf reduces false positive
names by ignoring affixes that can not form
names even if one possible solution of the stem is a name. 
Doing the same with Buckwalter and ElixirFM requires  
expanding the annotations resulting from one simple affix to be 
reflected in all the composite affixes that contain it.
%expanding the annotations resulting from one affix to reflect
%its multiple composing affixes.
%However, this was easily performed using our recursive definition 
%of the affixes.

@@

>>>>

%\title{Instructions for NAACL HLT 2010 Proceedings\Thanks{This...}}
\title{Application Specific Arabic Morphological Analyzer for Arabic Text Mining}

@@

\begin{document}
\maketitle
\begin{abstract}
Natural language processing (NLP) uses morphological 
analysis to abstract and annotate text.
Often times, 
%      and in particular in the context of the Arabic language, 
annotations resulting from 
a branch of morphological analysis may not be
appropriate for the case study at hand. 
In this paper we present Sarf, a case-based morphological 
analyzer for Arabic. 
Sarf allows a case-based controller to control and refine
the state of the analysis on the fly. 
It uses non-deterministic finite state Automata
to analyze a stream of text. 
%Each machine corresponds to a valid analysis. 
Sarf uses concatenative analysis based on recursive affixes 
to better preserve part of speech information.
We automated the analysis of three books of Islamic
narrations (hadith) using Sarf where we
abstracted each book into a vector of complex structures.
Our results show higher accuracy and better efficiency
compared to current analyzers.
\end{quote}
\end{abstract}

@@

%Automated analysis of Arabic data sets, including texts, 
%publications, records and digital media is essential
%with the huge digital Arabic content available nowadays. 
Morphological analysis is key in current automated 
analysis techniques for Arabic text. 
Current morphological analyzers~\cite{Sughaiyer:04}
use concatenative analysis when
considering the internal structure 
of an Arabic word and
composing it into several {\em morphemes}. 
A morpheme can be a {\em stem}, or an {\em affix}.
An affix can be a {\em prefix, suffix, } or an {\em infix}.
The analysis of one word may lead to several possible
morphological solutions.
\vocalize
For instance, the word \RL{'a.hmadH}~\footnote{In this document, we use the default ArabTeX transliteration style ZDMG.}
may have two valid morphological analyses. 
The letter \RL{'a} may be a prefix and the word means 
``I praise him'', or 
%The letter \RL{'a} may also be 
part of the stem \RL{'a.hmad} (a proper noun)
and the word means ``his Ahmad''.

\novocalize
The accuracy of the solutions suffer due to inherent difficulties
of morphological analysis of the Arabic language. 
For example, it is common practice to write Arabic text
without short vowels. 
This greatly increases the ambiguity of Arabic text. 
Arabic letters can have up to %(maybe reduce this sentence)
four different forms
corresponding to their position in a word, i.e, beginning,
middle, end of word and separate forms. 
This allows the phrase \transfalse
\RL{il_A\nospace almdrsT} \transtrue
to be visually recognizable
as two separate words \RL{il_A} (to) and \RL{almdrsT} (the school) 
without the need of a space in between. 
The reason is the first word \RL{il_A} ends with
\RL{_A} a non-connecting letter. 
These words,
referred to as ``run-on'' words~\cite{Buckwalter:04},
occur often, and greatly increase the
difficulty of tokenization.

{\bf Sarf.~~~}
In this paper, we present Sarf~\footnote{Sarf (name modified for blind review) 
has been submitted with the paper and will be available online as an open source tool.},
 a {\em novel case-based efficient
morphological analyzer} that uses 
non-deterministic finite state Automata 
driven by an application-specific controller.
Each machine in Sarf takes one letter at a time as input
and represents a valid morphological analysis.
The application-specific controller prunes false positives
early in the run by making a decision at every input letter
to control the machines.
The controller we present in Section~\ref{sec:controller}
is an FSA. However, as we explain in Section~\ref{sec:ndfsa},
it can be a logical entity of any type.

%Each alive machine in Sarf takes as input one letter at a time 
%from a text stream. 
Sarf does not assume that the word at hand is a token and
performs tokenization on the fly based on morphological correctness.
Up to our knowledge, this is the first morphological analyzer that 
handles the ``run-on'' words problem.

Sarf uses concatenative analysis with a novel 
{\em recursive affixes} refinement,
where an affix can be a sequence of other affixes which are concatenatively compatible.
%An atomic affix $a$ is a standalone affix
%that is not the concatenation of other affixes. 
%We define an affix $f: a | a f$ to be either
%an atomic affix $a$ or an atomic affix concatenated
%with an affix. 
Sarf introduces recursive affixes in order to
better retain part of speech information and enhance the 
efficiency of affix matching.

Our main focus when building Sarf, was not on improving the semantical rules and expanding the morphological alternatives, instead we worked on adding support for higher level control by the specific application utilizing the lower level morphological analyzer. In other words, application-specific controller, has full control over and directs the morphological analysis of Sarf. % (sentence should be reworded differently, but is needed to explain why we mention higher and lower accuracy hypothesis, which seems vague otherwise)

We hypothesize that many NLP case studies need the 
morphological analyzer to answer queries that do not need 
high accuracy at the low morphological analysis level.
The analysis at the higher levels 
(i.e. answering the application-specific query)
can often compensate for 
tolerable inaccuracy at lower morphological levels. 
For example, if a query concerns proper names and the 
analyzer is considering a 
prefix that connects only to a verb,
%the analyzer can answer without considering the 
%rest of the word.
we do not need to further analyze the rest of the word 
to provide a negative answer.

We find evidence to our hypothesis in~\cite{Maamouri:10}.
The addition of a new corpus to the Arabic Tree Bank 
(ATB)~\cite{Maamouri:04}
%required compliance with the DARPA GALE program 
%including several NLP tasks such as English translation 
%annotations and word-level alignment of Arabic and English. 
%The new data 
challenged the existing processing and 
annotation guidelines. 
This led to refining 
SAMA~\cite{Kulick:10} with 
alternative guidelines. 
This refinement was achieved only after the NLP 
task interacted with the morphological analyzer,
a flexibility our analyzer is designed to provide for the application-specific controller on the run. %(check if this sentence makes sense)
In addition, 
the work of Habash~\shortcite{Habash:06} reports that different 
types of morphological analyses behave differently on the same %case study. 
data set.
The analyzers may report different solutions or 
may present the solutions in different orders. 
Clearly, some of 
the presented solutions and their order of presentation may not be 
appropriate for the %case study.
application needs, which is an issue that Sarf tackles. % (maybe reworded)
Moreover, our results strongly support our claim that
an application-specific morphological analyzer that allows a application-specific 
controller to intervene at every decision to guide, use, order, and 
refine the morphological analysis is of high utility.
We validate our hypothesis using a case study from the Islamic 
literature, introduced later in Section~\ref{sec:islamic}.

The rest of this paper is structured as follows. In Section~\ref{sec:related}
we compare Sarf to related work. In Section~\ref{sec:sarf}
we discuss Sarf and illustrate it with a running example.
%We also explain how non-deterministic finite state Automata
%work. 
In Section~\ref{sec:islamic} we present the case
study. We present our results in Section~\ref{sec:results}
and conclude with future work in Section~\ref{sec:future}.

@@

@@

The survey in~\cite{Sughaiyer:04} compares
several morphological analyzers. 
Analyzers such as~\cite{Khoja:01,Darwish:02} 
target specific applications in the 
analyzer itself or use a specific set of POS tags
as their reference.
Sarf differs in that it is a general morphological 
analyzer that reports all possible solutions. 
It is application-specific in the sense that a controller 
prunes these solutions.

Current morphological analyzers such as 
Buckwalter~\shortcite{Buckwalter:02},
Beesley~\shortcite{Beesley:01},
SAMA~\cite{Kulick:10},
and ElixirFM~\cite{Otakar:07} 
take as input white space delimited tokens~\cite{Kulick:10},
consider them as words,
and enumerate all possible solutions. 
%This approach has several problems. 
However, we note the following points that deserve consideration:
%(We need not to mention especially point 2 as a problem according to reviewer comments)
First, a white space delimited token may have 
more than one word.
%Second, the exhaustive enumeration may hurt performance and may
%not be necessary or appropriate
%in some case studies as noted in~\cite{Maamouri:10}. 
Second, in many applications, the exhaustive enumeration may not be needed or appropriate as noted in~\cite{Maamouri:10}; consequently, adding unnecessary running time.
Other morphological analyzers such as 
Amira~\cite{Diab:07,Benajiba:07},
MAGEAD~\cite{Habash:05}, and MADA+TOKAN~\cite{Habash:09} 
% (reviewer wants to distinguish MADA as a tool for morphological disambiguation instead of analysis)
use machine learning and support vector machines (SVM) 
to enhance the accuracy of the morphological analysis at the expense 
of performance. % (we should stress that we mean performance in time and not accuracy although evident since some reviewers did not get)
Xerox rules can be compiled into specialized finite state
machine (FSM) based analyzers as described in~\cite{Beesley:03}.
However, the efficiency of the resulting analyzer depends on the
way the Xerox rules are written. 
This requires deep knowledge and insight from the user
in compilation techniques, context free grammars, 
and morphological analysis.
%context free grammar rules that can be automatically compiled
%into efficient non-deterministic finite state Automata (NDFSA).

Sarf builds upon the lexicon of Buckwalter\shortcite{Buckwalter:02}.
It differs from Buckwalter in that it defines a shorter list of affixes
and a longer list of affix compatibility rules to allow 
affixes to be defined recursively
so that we can have prefix-prefix an suffix-suffix 
concatenations.
This allows Sarf to better maintain and propagate 
the POS tags associated with affixes. 
%This information is highly 
%redundant with Buckwalter.% and that may lead to consistency issues. 
Sarf keeps all the possible analyses alive where each analysis
corresponds to a set of decisions and positions in the text. 
Buckwalter produces a set of solutions for each token 
and the several analysis threads of the text need to be 
generated as a product of the token solutions.

Similar to Sarf, Beesley~\shortcite{Beesley:01} uses
finite state Automata (FSA). 
It reports that the number of Automata generated by a compiler
for Xerox rules can not be controlled and also reports a difficulty 
to compose the FSAs into a single framework. 
Sarf however constructs an NDFSA from
the controller and the prefix, stem, and suffix deterministic 
transducers to solve the problem.
Sarf also differs in that it uses recursive affixes in the 
affix transducers, and in that the transducers consider
input from the controller to modify their behavior. 
%With Sarf it is easy to control the FSA with a case-based
%controller. 
Beesley requires recompilation of the Xerox rules. % (don't we do also for any change in the controller code?)

@@

Our approach is close to the refinement of SAMA~\cite{Maamouri:10}.
SAMA was refined to interact with
the ATB~\cite{Maamouri:04} project after the addition of a large 
new corpus. 
%We find supporting evidence in this decision to our hypothesis
%about the utility of a case-based analyzer. 
The algorithmic changes in SAMA were
done manually and worked in integration with the ATB format. 
Sarf is an application-specific controller approach that can modify 
the behavior of the analyzer on the fly and can interact
with any case study. 
%In Sarf, the analyzer remains general, and the 
%case-based controller specializes the analysis.

@@

Like ElixirFM~\cite{Otakar:07}, Sarf builds on the lexicon
of the Buckwalter analyzer. 
Sarf also uses deterministic parsing with tries 
to implement the affix and stem transducers. 
We think that the inferential-realizational approach 
of ElixirFM
that is highly compatible with the Arabic linguistic 
description~\cite{Badawi:04}
can benefit from many features unique to the Arabic language.
Sarf leaves implementing that to the application-specific controller
since in many cases the abstraction needs only a partial 
linguistic model of the Arabic language. 
%We can compare the case-based controller in Sarf to a high level
%component in the ElixirFM domain specific language.

@@

Before moving with the letter \RL{b} to the accept state,
$\Phi$ needs to make sure that the stem \RL{l`b} is compatible
with the prefix \RL{wsy}. 
Sarf keeps compatibility category values as part
of the accept states. 
%Sarf considers the value of the category as part of the state.
Thus each accept state in Figure~\ref{f:example} represents
more than one concrete state. 
If the category of \RL{l`b} is compatible with the category of
\RL{wsy} then \RL{b} will move $\Phi$ to an accept state. 
Otherwise, it will move it to a regular or a reject state.

Since $\Phi$ is now in an accept state in ${\cal S}$, it 
spawns $\Xi$ which transitions to the root of ${\cal X}$. 
After it consumes \RL{h--} $\Phi$ dies since there is no
\noTrRL{h--}-edge from the current state.
$\Xi$ represents a valid analysis and proceeds.

@@

% TODO: check how to merge this part here
\transfalse
\begin{figure}[tb]
\center{
\resizebox{.9\columnwidth}{!}
{ \input{figs/exhadith.pdftex_t}}
\caption{Hadith abstraction example.}
\label{f:exhadith}
}
\end{figure}
% the FSMs for the three words
\transtrue

We validate our hypothesis using a case study from the Islamic 
literature. 
A \RL{.hady_t} is a narration related to the prophet Mohammad
through a \RL{sanad} or a sequence of narrators. 
Figure~\ref{f:exhadith} shows an example \noArRL{.hady_t} in 
Arabic with its transliteration and translation. 
The authenticity of a \noArRL{.hady_t} depends on 
the credibility of the narrators as reported in 
separate biography books. 
In this paper, we consider the problem of automatically segmenting
a \noArRL{.hady_t} book into narrations, each 
narration into its \RL{matn} (content) and its \noArRL{sanad},
and each sanad into the 
separate narrators. Later, we can look each narrator 
up in the biography books. 
For example, 
Figure~\ref{f:exhadith} shows proper names in boxes connected
to form complex names of narrators. 
For example, 
\noTrRL{qtybT} is the first name 
of narrator $n_1$, and 
\noTrRL{s`yd} is the name of his father as 
the word \noTrRL{bn} (son of) indicates.

The hadith case study is interesting to Sarf since the controller
considers %most of the analyses where we have a concentration
proper names and ignores text where
proper names occur rarely.% as long %as the text is valid.
We are also interested in discovering words 
that relate names to each other such as \RL{bin} or ``the son of''
and words that mean narrate such as \RL{qaal} or ``said''.

With Sarf, we successfully automated the analysis of 
three books of Islamic narrations~\cite{IbnHanbal,AlTousi,AlKulayni}
and extracted from each one of them a vector of complex structures 
that are three level deep in hierarchy.
%We also showcase that the results of the accurate high level 
%abstraction
%results of the case-based analysis was not affected by a major
%refinement to the analyzer that increased low level analysis accuracy.

A hadith in Islamic literature is a narration from the prophet Mohammad
related by multiple narrators.
Establishing the authenticity of a hadith is an important task
in Islamic studies. 
Based on them Islamic scholars issue rules that affect the life
of people around the world. 
The literature lacks an exhaustive mechanism of authentication
checks. This is currently manual work prone to error and almost
impossible to complete in the life time of a scholar due to the
large literature available. 
%Several researchers~\cite{Hadithopaedia:08} attempted to automate 
%the analysis of the hadith literature. 
In this paper, we use Sarf to successfully automate the
analysis of three books of hadith selected 
arbitrarily~\cite{IbnHanbal,AlKulayni,AlTousi}\footnote{We obtained
  the digitized books from online sources such as 
  \href{http://www.yasoob.com/}{http://www.yasoob.com/} and 
  \href{http://www.al-eman.com/}{http://www.al-eman.com/}. }
%  Our understanding is that those books are historical documents with 
 % public IP.}.

@@

We investigated the case of Al-Kafi where the accuracy of 
the simple analysis was significantly lower than that of the 
refined analysis. 
We discovered that the chains detected by the refined analysis 
were a subset of those detected by the simple
analysis. 
The reason for the lower accuracy is false positive chains, 
implying that the simple analysis did not miss important 
information. 
We hypothesize that a higher level of abstraction such as 
intersecting narrators from the hadith books with those from
the biographies can reduce the false positives in the chains.

%\subsection{Buckwalter and ElixirFM}
{\bf Buckwalter and ElixirFM.~~~}
We implemented our case study with Buckwalter in two phases. 
In phase one, we processed the entire text with Buckwalter and
obtained annotations to text. 
In the second phase, we fed the annotations to 
the hadith controller as if they were coming from Sarf. 
We measured the time it took to accomplish both tasks
ignoring the time spent streaming the annotations.
We did the same for ElixirFM.

Sarf outperforms both Buckwalter and ElixirFM in terms of 
running time. 
It also produces higher accuracy measures 
mainly due to the extended
lexicon as well as to the fact that Sarf reduces false positive
names by ignoring affixes that can not form
names even if one possible solution of the stem is a name. 
%Doing the same with Buckwalter and ElixirFM requires  
%expanding the annotations resulting from one simple affix to be 
%reflected in all the composite affixes that contain it.
%expanding the annotations resulting from one affix to reflect
%its multiple composing affixes.
%However, this was easily performed using our recursive definition 
%of the affixes.

@@

%TODO: merge into conclusion, taken from introduction
We make the following key contributions: 
\begin{enumerate}
\item {\bf Case-based analysis:}  We designed and implemented
a case-based morphological analyzer where a controller
guides the analysis. 
%\item {\bf Exhaustive analysis:} Since we have a case-based 
%controller that can decide on the fly whether an analysis is 
%accepted or not, we can afford to keep all valid analyses modulo
%those pruned by the case controller. 
%We do that using non-deterministic FSAs. 
\item {\bf Recursive affixes:}
We refine the concatenative analysis of 
Buckwalter~\shortcite{Buckwalter:02} and define affixes
recursively. 
This allows Sarf to perform better and 
maintain compositional part of speech (POS) tags.
\item {\bf Hadith case study:}
%We evaluated Sarf using the hadith case study. 
We were able to efficiently extract a complex three 
level hierarchical 
structure containing the chains of narrators 
with high accuracy. 
\end{enumerate}
